What We Eval in the Shadows

### 2021-08-10 paper/main.tex
Add miscellaneous section

\mypara{Domain-Specific Languages} The combination of laziness, metaprogramming,
first-class environments and dynamic evaluation makes R the perfect toolkit for
building DSLs. An example of this is the \code{glue} package which provides
functions for string interpolation. The code snippet below shows an example of
\code{glue} function in action.

@@

\mypara{Boilerplate Removal} The combination of \code{match.call} and \code{eval}
is frequently used by modeling functions to avoid boilerplate code and ease
maintenance. Consider the example shown below of function \code{survfit.formula}
from the \code{survival} package.

@@

\paragraph{Miscellaneous}

\paragraph{Lookup and indexing}

For example, the \code{PopED::get\_rse} uses it to
assign a number of variables in the current scope:
\code{eval(parse(text=paste(default_args[[i]]), "<- ", i))}. There are also more
sophisticated cases. The \code{plyr::allocate\_column} function uses eval for a
deferred assignment to fill missing values in a column. Package overture
repeatedly evaluates R expressions capturing any assignments that happen in it
as samples of Markov chains.

>>>>

\paragraph{Domain-Specific Languages} The combination of laziness, metaprogramming,
first-class environments and dynamic evaluation makes R the perfect toolkit for
building DSLs. An example of this is the \code{glue} package which provides
functions for string interpolation. The code snippet below shows an example of
\code{glue} function in action.

@@

\paragraph{Boilerplate Removal} The combination of \code{match.call} and \code{eval}
is frequently used by modeling functions to avoid boilerplate code and ease
maintenance. Consider the example shown below of function \code{survfit.formula}
from the \code{survival} package.

@@

\paragraph{Miscellaneous} Many uses of \eval are diverse and do not fit into a
category by themselves. For example, packages for unit testing (testthat,
testit, etc.), code benchmarking (rbenchmark, microbenchmark, etc.), and,
running code in parallel (foreach, doParallel, etc.), use \eval to evaluate
program fragments in a particular way in specific environments. Many numerical
computation and mathematical modeling packages use \eval to evaluate chunks of
code or bodies of functions in specific environments. Consider the
\code{findZeros} function from \code{mosaic} package which is used to compute
the zero of a function.

\begin{figure}[h]
\begin{lstlisting}
findZeros <- function(expr, ..., xlim=c(near-within, near+within), near=0, within=Inf) {
  dots <- list(...)

  pfun <- function(x){  # removed . from name, was .x
    mydots <- dots
    mydots[[rhsVars]] <- x
    eval( lhs(expr), envir=mydots, enclos=parent.frame() )
  }
  searchx <- sort(unique(internal.near   + c(0, leftseq, middleseq, rightseq)))
  y <- sapply( searchx, pfun )
\end{lstlisting}
  \caption{mosaic::findZeros}
\end{figure}

This function can be invoked as: \code{findZeros( A*sin(2*pi*t/P) ~ t,
  xlim=c(0,100), P=50, A=2)}. It will evaluate \code{A*sin(2*pi*t/P)} inside
\code{pfun}. It calls \code{pfun} repeatedly with different values of \code{t}
to find the point at which the expression is zero.
